---
title: "STAT380 Final Project"
author: "Jacob Gavin, Bryson Davis, Abhilash Katigiri, Taegwon Lee"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# Load libraries
library(readr)
```

```{r}
# Import datasets
games_p1 <- read.csv('./data/CODGames_p1_380.csv')
games_p2 <- read.csv('./data/CODGames_p2_380.csv')
gamemodes <- read.csv('./data/CODGameModes.csv')
maps <- read.csv('./data/CODMaps.csv')

```

# Data Cleaning and Visualization (no GenAI)
Relevant Information: (Complete without using Generative AI) Prior to each online match, players in the
game lobby are presented with two options for the battlefield of the upcoming game (`Map1` and
`Map2`). The players have the option to vote and the resulting vote is recorded in the `MapVote` column.
The winning map is listed in the `Choice` column. In the event of a tie vote, the map listed in `Map1` is
chosen. (Games for which the player entered the lobby after the vote has taken place have no information
in `Map1` and `Map2` but have the winning map presented in `Choice`.)
4
Research Question: Which maps are the most likely to win the map vote when they are an option?
Notes: To answer this question, write a paragraph (or more) discussing how you plan to answer this
question. Be sure to address the data quality issues mentioned below and discuss how you will do the
calculations. Then, write code and answer the question. (If I must answer questions about your
approach/decision making process by reading your code rather than your discussion, you will lose points.)
As part of your solution, you should calculate the proportion/probability that a map wins the vote given
that it was a candidate. To do this, you will have to calculate the number times that each map was listed as
a candidate (Map1 or Map2) and earned more votes than the other candidate. As part of this, you should
consider whether a given map won the vote by getting more votes than the other option or if it was
selected since it was `Map1` and the vote was a tie. You should also include a visualization of the results.
There might be some data quality issues (such as misspelled map names and extra (trailing) blanks in
some entries) to solve for this problem. You can find the proper names/spellings in the CODMaps.csv
file. To full receive full credit, you must write code to solve these issues rather than editing the .csv files

# Data Cleaning and Visualization (with GenAI)
Repeat Task 1 using a generative AI of your choice. To answer this question, mention the tool (including
version number if appropriate) you have selected. Then, discuss the prompt(s) you have used and provide
the solution produced by the generative AI. While it is fine to paste the question into the generative AI as
your first prompt, you should also use additional follow-up prompts if it is beneficial to do so. Be sure to
discuss all prompts used in your report.
Then, implement the generative AI solution.
Finally, and most importantly, you should compare your solution from Task 1 to the generative AI
solution. Discuss similarities/differences, strengths/weaknesses, etc., and provide an overall assessment of
which solution is better. The discussion should consider the correctness of the answers and should be
substantial. Demonstrate that you have given the comparison considerable thought by making at least 3
substantial points as part of your comparison. Each point should take the form of a well-written
paragraph.

# Inference

Relevant Information: There are a variety games types (GameType variable) within this dataset. The
difference between the game types is that players have different objectives for the game. For instance, in
the game type “Hardpoint”, teams earn points by capturing and defending a location. In “TDM” teams
earn points by eliminating enemy opponents. As these game types have different objectives and may last
for different amounts of time, the game type might affect the TotalXP earned.
Research Question: How does the game type affect TotalXP after accounting for the Score?

Notes: Score refers to the player’s score, not the “score” of the match (i.e., not the Result column). This
answer requires some data wrangling that may require knowledge that we have not covered. (Again, part
of the skillset you are working to develop is learning how to answer questions you have not seen
previously.) In particular, there is no distinction between HC – TDM and TDM, no difference between
HC – Hardpoint and Hardpoint, and so on for the other game types. Write code to clean the values in the
GameType column to reflect this information. Then, perform an exploratory data analysis by create
appropriate visualizations/summary statistics that explore the distribution of the variables and show the
relationship between TotalXP, Score, and GameType. (You decide on the type/number of visualizations,
but the analysis should be complete.) Finally, build an appropriate model for TotalXP based on Score and
GameType. You should use the model to then answer the research question.



# Prediction

Relevant Information: In this task, your goal is to compare a variety of classification methods. In
particular, you should write your own research question that can be answered by comparing the
effectiveness of various classification methodologies. To demonstrate your understanding of these
methods, you should implement two classification methods from class, one of which must be random
forest, and a third method that we will not cover in class. (The purpose of the using a method we did not
cover is I want you to practice learning about a method and its implementation on your own. Basically,
find a tutorial that explains the method and how to implement it.) You will then have to compare the
results and decide which method was the most effective.
Research Question: Write your own question and be sure that the question and answer are clearly written
in your report.
Notes: Since you will be using random forest, do not use a decision/classification tree as one of your other
methods. For this problem, you should provide a brief description of the methods that you will use. (A
description is more than listing the name of the procedure. You should describe how the procedure
works.) You will implement and compare the effectiveness of these methods. As part of this process, you
will have to make a number of decisions such as whether you will do any data wrangling (maybe you
remove partial matches, maybe you create new variables, etc.), which methods will you use, how will you
fairly compare the results between methods, which method is best etc. All of these decisions should be
included in your report. If I have to learn about your decisions/analysis by reading your code, you will
lose points.





